---
title: "Correia Consultant Agency"
author: "Pedro Morais, Braedon Fiume, Matt Adams"
date: "Add a date"
output:
  rmdformats::readthedown:
    css: "custom.css"
    thumbnails: false
    highlight: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
library(tidyverse)
library(rvest)
library(knitr)
library(janitor)
library(tidytext)
library(countrycode)
library(rworldmap)
library(viridis)
library(GGally)
library(plotly)
library(dplyr)
library(stringr)
library(leaflet)
library(gganimate)
library(lubridate)
library(maps)
library(ggthemes)
library(RColorBrewer)

starbucks <- read_csv("Data/starbucks.csv")
city_income <- read_csv("Data/kaggle_income.csv")
county_pop <- read_csv("Data/us_county.csv")
unemployment <- read_csv("Data/unemployment.csv")
starbucks_positive <- read_csv("Data/starbucks_positive.csv")
starbucks_negative <- read_csv("Data/starbucks_negative.csv")

#Standard unemployment data used for each county
unemployment_county <- unemployment %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>%
  group_by(County) %>%
  na.omit() %>%
  summarize("Avg. Unemployment Rate" = mean(Rate, na.rm=TRUE)) 

```

# Introduction
Starbucks’ gargantuan $26 billion of revenue in 2019 is resounding evidence that Americans, above everything, love their coffee. They prove it time and time again, most recently during the current pandemic: some, unable to relinquish their daily ‘bucks, waited hours on end in long drive-thru lines to get their fix when coffee shops were closed for in-person operations.
	
It is no surprise, then, that Starbucks owns more than 8,000 stores across the US, and continues to grow everyday. These sheer numbers are reflective of a company that is about more than just coffee, having a significant influence on US society and culture with (often controversial) initiatives such as removing religious references from holiday-themed cups. In addition, Starbucks is known for treating their employees extremely well by providing health coverage, tuition coverage, and 401(k) plans, which again reinforces their highly-regarded brand.

As such, the astounding popularity of the chain in the country and the prospect of further expansion raises important questions. What can the current distribution of Starbucks stores tell us about societal factors across the country? Which factors should the brand consider when expanding into new locations? How can companies like Starbucks, which pride themselves in a positive social and environmental outlook, incorporate such values into their corporate strategy - especially in the current pandemic?

## Distribution of Starbucks Across the World and US

```{r figure1, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

#World Map of Starbucks
worldmap <- getMap(resolution = "coarse")
plot(worldmap, bg = "white", col = "black")
points(starbucks$Longitude, starbucks$Latitude, 
       col = "#006d2c", cex = .01)

```

Text explaining world map

```{r figure2, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

#US locations; except Hawaii and Alaska
us_stars <- starbucks %>%
  filter(Country=="US") %>%
  rename(State=`State/Province`) %>%
  filter(State!="AK") %>%
  filter(State!="HI")
  
usa <- map_data(map = "state"
                       , region = ".")
usa_states <- usa %>%
  select(region, group) %>%
  group_by(region) %>%
  summarize(
    group=sum(group)/n()
  )

state_info <- data.frame(state_full = tolower(state.name)
                         , State = state.abb
                         , Region = state.region)

us_map_star <- us_stars %>%
  left_join(state_info, by = "State") %>%
  left_join(usa_states, by = c("state_full" = "region"))

us <- map_data(map = "state"
                       , region = ".") 

usa_states <- map_data(map = "state"
                       , region = ".")

city_star <- us_map_star %>%
  group_by(City) %>%
  summarize(
    Count = n()
  )

city_count_star <- us_map_star %>%
  left_join(city_star, by = "City") %>%
  mutate(Count = round(Count, 1))

#STARBUCKS LOCATIONS IN THE US
ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="gray", alpha=0.3) +
  geom_point(data=city_count_star, aes(x = Longitude, y = Latitude, group=group), color = "#006d2c", alpha = 0.5) +
  theme_void() +
  scale_size(guide = "none") 

```

TEXT EXPLAINING INTRODUCTION VISUAL OF STARBUCKS LOCATIONS

Maybe include magnitude visualization, explain how it helped our understanding of how current starbucks locations are in clusters 

Process Towards Selecting New Location:
1. Find States
2. Find Counties

#Determining States of New Starbucks Locations

Paragraph about how we went about determining clusters of states, and explaining why we want to use clusters and explaining factors we use

## State-Level Factors

- Consumer Sentiments
- Average Income
- Average Unemployment Rate
- Number of Starbucks

##Cosumer Sentiments Across US

Explain how we collected data through twitter and used colleciton of words to find proportion of positive and negative words of tweets containing the word "starbucks" and had a geo-location, used 18,000 tweets.


```{r figure3, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

ggplot(starbucks_positive, aes(x = long, y = lat, group = group
                      , fill = positive.proportion)) +
  scale_fill_viridis() +
  geom_polygon(color = "white") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(fill = "Proportion of Positive Words in Tweets Per State",
       caption = "*Grey states have no twitter data") +
    theme(legend.title = element_text(color = "black", size = 8),
        legend.text = element_text(color = "black", size = 8))

```


```{r figure4, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

ggplot(starbucks_negative, aes(x = long, y = lat, group = group
                      , fill = negative.proportion)) +
  scale_fill_viridis() +
  geom_polygon(color = "white") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(fill = "Proportion of Negative Words in Tweets Per State",
       caption = "*Grey states have no twitter data") +
    theme(legend.title = element_text(color = "black", size = 8),
        legend.text = element_text(color = "black", size = 8))

```

##Income and Unemployment Data

Explain data came from a kaggle and a county level, which we summarized into states


##State-Level Clusters

Explain cluster analysis and elbow plot helped to determine 5 clusters were needed


```{r figure5, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

income_state <- city_income %>%
  group_by(State_Name, State_ab) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) %>%
  mutate(State = tolower(State_Name)) %>%
  select(State, `Avg. Income`, State_ab)

unemployment_state <- unemployment %>%
  group_by(State) %>%
  na.omit() %>%
  summarize("Avg. Unemployment Rate" = mean(Rate, na.rm=TRUE)) %>%
  mutate(State = tolower(State))

starbucks_state <- starbucks %>%
  group_by(`State/Province`) %>%
  na.omit() %>%
  summarize("Num. Starbucks" = n())
  
star_neg <- starbucks_negative %>%
  select(negative.proportion, state_full) %>%
  group_by(state_full) %>%
  na.omit() %>%
  summarize(
    negative.proportion = mean(negative.proportion, na.rm=TRUE)
  )

state_factors <- income_state %>%
  left_join(unemployment_state, by="State") %>%
  left_join(star_neg, by = c("State"="state_full")) %>%
  left_join(starbucks_state, by = c("State_ab" = "State/Province"))

state_factors$`Avg. Unemployment Rate`[10] <- 7.6

final_state_factors <- state_factors %>%
  na.omit() %>%
  ungroup()

final_final_state_factors <- final_state_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) 

only_std_values <- final_final_state_factors %>%
  select(`Avg. Unemployment Rate_std`, `Avg. Income_std`, negative.proportion_std, `Num. Starbucks_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=10, ncol=2)

#Collecting information for each cluster 
for (i in 1:10){
  fig[i,1] <- i
  fig[i,2] <- kmeans(only_std_values, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:10)) +
  labs(x = "K", y = expression("Total W"[k]), title="Cluster Analysis")

```

###Visualization of Cluster
So we chose 5 clusters based on this plot, and the 5 clusters are shown on this visualization:


```{r figure6, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

km_out_std <- kmeans(only_std_values, centers = 5, nstart = 20)

clust_5_std <- final_final_state_factors %>%
  mutate(clust5_std = as.character(km_out_std$cluster)) %>%
  arrange(clust5_std) %>%
  mutate(State_Name=tolower(State_Name))

usa_states <- map_data(map = "state"
                       , region = ".")

final_cluster <- clust_5_std %>%
  left_join(usa_states, by = c("State_Name" = "region"))

ggplot() +
geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
geom_polygon(data= final_cluster, aes(x = long, y = lat, group = group, fill = clust5_std), color="black") +
  geom_polygon(color = "black") +
  theme_void() +
  coord_fixed(ratio = 1.3) +
  labs(fill = "5 Clusters of States") +
  theme(legend.position="bottom")

```

###Cluster Characteristics
Explain we can see characteristics of each cluster through this matrix. Explain each cluster's unique properties

```{r figure7, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vars_std = c("Avg. Unemployment Rate_std", "Avg. Income_std", "negative.proportion_std", "Num. Starbucks_std")

ggpairs(data = final_cluster
        , aes(color = clust5_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))


```

##Final List of States

Then, we calculated the center within each center, finding the states within each cluster that best represented each cluster by finding the state with the miminal distance between each characterstic's value and itself. From this analysis we concluded the final 5 states, each representative of their own cluster, are: 


```{r figure8, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

#Code we used for each cluster to determine its' representative state
clust_1 <- final_state_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  mutate(clust3_std = as.character(km_out_std$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(State_Name=tolower(State_Name)) %>%
  filter(clust3_std==1) 

clust_1 <- clust_1 %>%
  mutate(dist = sqrt(
                      (`Avg. Unemployment Rate_std` - km_out_std$centers[2,1])^2 +
                      (`Avg. Income_std` - km_out_std$centers[2,2])^2 +
                      (`negative.proportion_std` - km_out_std$centers[2,3])^2 +
                      (`Num. Starbucks_std` - km_out_std$centers[2,3])^2)
  ) %>%
  arrange(dist)


```

- Virginia
- Vermont
- Nevada
- Wisconsin
- California

# Determining Counties of New Starbucks Locations

##County-Level Factors

- Average Income
- Average Unemployment Rate
- Number of Starbucks

Step-by Step Process

1. Cluster Analysis
2. Creating clusters based on appropriate number of centers
3. Determine the most suitable cluster of counties

##Clustering at a County Level For Each Selected State

###Virginia

Cluster Analysis
```{r figure9, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

va_income_city <- city_income %>%
  filter(State_Name=="Virginia") %>%
  group_by(City) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) 

va_starbucks_city <- starbucks %>%
  filter(`State/Province`=="VA") %>%
  group_by(`City`) %>%
  summarize("Num. Starbucks" = n())

va_city_info <- va_income_city %>%
  left_join(va_starbucks_city, by = c("City")) %>%
  mutate_all(~replace(., is.na(.), 0))

va_county_info <- city_income %>%
  filter(State_ab=="VA") %>%
  select(City, County) %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>% 
  distinct() %>%
  left_join(va_city_info, by = c("City")) %>%
  group_by(County) %>%
  summarize(
    "Avg. Income" = mean(`Avg. Income`), 
    "Num. Starbucks" = sum(`Num. Starbucks`)
  )

va_county_factors <- va_county_info %>%
  left_join(unemployment_county, by = c("County")) %>%
  na.omit()

va_final_county <- va_county_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  select(County, `Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)
  
va_final_county_factors <- va_final_county %>%
  select(`Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=10, ncol=2)

#Collecting information for each cluster 
for (i in 1:10){
  fig[i,1] <- i
  fig[i,2] <- kmeans(va_final_county_factors, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:10)) +
  labs(x = "K", y = expression("Total W"[k]))

```


#Visualizing clusters

```{r figure10, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(75)
va_county_km <- kmeans(va_final_county_factors, centers = 5, nstart = 20)

# add cluster assignments to the data frame
va_county_clust <- va_final_county %>%
  mutate(clust3_std = as.character(va_county_km$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(County=tolower(County))

usa_states <- map_data(map = "county"
                       , region = "virginia") %>%
  select(subregion, long, lat, group)

va_final_cluster <- va_county_clust %>%
  mutate(County=str_trim(County)) %>%
  left_join(usa_states, by = c("County" = "subregion"))

ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
  geom_polygon(data=va_final_cluster, aes(x = long, y = lat, group=group, fill=clust3_std), color="black") +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  labs(title = "Clusters in Tennessee",
       caption = "*Grey states have no income or unemployment information",
       fill="Clusters") +
  theme_void()

```


Cluster characteristics, explain why we chose cluster 4

```{r figure11, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vars_std = c("Avg. Income_std", "Num. Starbucks_std", "Avg. Unemployment Rate_std")

ggpairs(data = va_county_clust
        , aes(color = clust3_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))

#Compile for the end result
va_poor_county <- va_county_clust %>%
  filter(clust3_std==4) %>%
  mutate(County=str_to_title(County)) %>%
  inner_join(va_county_factors, by="County") %>%
  select(County, `Avg. Income`, `Num. Starbucks`,`Avg. Unemployment Rate`)

```


##Vermont

Cluster Analysis
```{r figure12, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vt_income_city <- city_income %>%
  filter(State_Name=="Vermont") %>%
  group_by(City) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) 

vt_starbucks_city <- starbucks %>%
  filter(`State/Province`=="VT") %>%
  group_by(`City`) %>%
  summarize("Num. Starbucks" = n())

vt_city_info <- vt_income_city %>%
  left_join(vt_starbucks_city, by = c("City")) %>%
  mutate_all(~replace(., is.na(.), 0))

vt_county_info <- city_income %>%
  filter(State_ab=="VT") %>%
  select(City, County) %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>% 
  distinct() %>%
  left_join(vt_city_info, by = c("City")) %>%
  group_by(County) %>%
  summarize(
    "Avg. Income" = mean(`Avg. Income`), 
    "Num. Starbucks" = sum(`Num. Starbucks`)
  )

vt_county_factors <- vt_county_info %>%
  left_join(unemployment_county, by = c("County")) %>%
  na.omit()

vt_final_county <- vt_county_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  select(County, `Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)
  
vt_final_county_factors <- vt_final_county %>%
  select(`Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=7, ncol=2)

#Collecting information for each cluster 
for (i in 1:7){
  fig[i,1] <- i
  fig[i,2] <- kmeans(vt_final_county_factors, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:7)) +
  labs(x = "K", y = expression("Total W"[k]))

```


#Visualizing clusters

```{r figure13, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(75)
vt_county_km <- kmeans(vt_final_county_factors, centers = 4, nstart = 20)

# add cluster assignments to the data frame
vt_county_clust <- vt_final_county %>%
  mutate(clust3_std = as.character(vt_county_km$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(County=tolower(County))

usa_states <- map_data(map = "county"
                       , region = "vermont") %>%
  select(subregion, long, lat, group)

vt_final_cluster <- vt_county_clust %>%
  mutate(County=str_trim(County)) %>%
  left_join(usa_states, by = c("County" = "subregion"))

ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
  geom_polygon(data=vt_final_cluster, aes(x = long, y = lat, group=group, fill=clust3_std), color="black") +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  labs(title = "Clusters in Tennessee",
       caption = "*Grey states have no income or unemployment information",
       fill="Clusters") +
  theme_void()

```

Cluster characteristics, explain why we chose cluster 1

```{r figure14, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vars_std = c("Avg. Income_std", "Num. Starbucks_std", "Avg. Unemployment Rate_std")

ggpairs(data = vt_county_clust
        , aes(color = clust3_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))

vt_poor_county <- vt_county_clust %>%
  filter(clust3_std==1) %>%
  mutate(County=str_to_title(County)) %>%
  inner_join(vt_county_factors, by="County") %>%
  select(County, `Avg. Income`, `Num. Starbucks`,`Avg. Unemployment Rate`)

```

##Nevada

Cluster Analysis
```{r figure15, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

nv_income_city <- city_income %>%
  filter(State_Name=="Nevada") %>%
  group_by(City) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) 

nv_starbucks_city <- starbucks %>%
  filter(`State/Province`=="NV") %>%
  group_by(`City`) %>%
  summarize("Num. Starbucks" = n())

nv_city_info <- nv_income_city %>%
  left_join(nv_starbucks_city, by = c("City")) %>%
  mutate_all(~replace(., is.na(.), 0))

nv_county_info <- city_income %>%
  filter(State_ab=="NV") %>%
  select(City, County) %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>% 
  distinct() %>%
  left_join(nv_city_info, by = c("City")) %>%
  group_by(County) %>%
  summarize(
    "Avg. Income" = mean(`Avg. Income`), 
    "Num. Starbucks" = sum(`Num. Starbucks`)
  )

nv_county_factors <- nv_county_info %>%
  left_join(unemployment_county, by = c("County")) %>%
  na.omit()

nv_final_county <- nv_county_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  select(County, `Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)
  
nv_final_county_factors <- nv_final_county %>%
  select(`Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=8, ncol=2)

#Collecting information for each cluster 
for (i in 1:8){
  fig[i,1] <- i
  fig[i,2] <- kmeans(nv_final_county_factors, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:8)) +
  labs(x = "K", y = expression("Total W"[k]))


```


#Visualizing clusters

```{r figure16, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(75)
nv_county_km <- kmeans(nv_final_county_factors, centers = 4, nstart = 20)

# add cluster assignments to the data frame
nv_county_clust <- nv_final_county %>%
  mutate(clust3_std = as.character(nv_county_km$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(County=tolower(County))

usa_states <- map_data(map = "county"
                       , region = "nevada") %>%
  select(subregion, long, lat, group)

nv_final_cluster <- nv_county_clust %>%
  mutate(County=str_trim(County)) %>%
  left_join(usa_states, by = c("County" = "subregion"))

ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
  geom_polygon(data=nv_final_cluster, aes(x = long, y = lat, group=group, fill=clust3_std), color="black") +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  labs(title = "Clusters in Tennessee",
       caption = "*Grey states have no income or unemployment information") +
  theme_void()

```

Cluster characteristics, explain why we chose cluster 4

```{r figure17, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}


vars_std = c("Avg. Income_std", "Num. Starbucks_std", "Avg. Unemployment Rate_std")

ggpairs(data = nv_county_clust
        , aes(color = clust3_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))

nv_poor_county <- nv_county_clust %>%
  filter(clust3_std==4) %>%
  mutate(County=str_to_title(County)) %>%
  inner_join(nv_county_factors, by="County") %>%
  select(County, `Avg. Income`, `Num. Starbucks`,`Avg. Unemployment Rate`)

```

##Wisconsin

Cluster Analysis
```{r figure18, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

wi_income_city <- city_income %>%
  filter(State_Name=="Wisconsin") %>%
  group_by(City) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) 

wi_starbucks_city <- starbucks %>%
  filter(`State/Province`=="WI") %>%
  group_by(`City`) %>%
  summarize("Num. Starbucks" = n())

wi_city_info <- wi_income_city %>%
  left_join(wi_starbucks_city, by = c("City")) %>%
  mutate_all(~replace(., is.na(.), 0))

wi_county_info <- city_income %>%
  filter(State_ab=="WI") %>%
  select(City, County) %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>% 
  distinct() %>%
  left_join(wi_city_info, by = c("City")) %>%
  group_by(County) %>%
  summarize(
    "Avg. Income" = mean(`Avg. Income`), 
    "Num. Starbucks" = sum(`Num. Starbucks`)
  )

wi_county_factors <- wi_county_info %>%
  left_join(unemployment_county, by = c("County")) %>%
  na.omit()

wi_final_county <- wi_county_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  select(County, `Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)
  
wi_final_county_factors <- wi_final_county %>%
  select(`Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=10, ncol=2)

#Collecting information for each cluster 
for (i in 1:10){
  fig[i,1] <- i
  fig[i,2] <- kmeans(wi_final_county_factors, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:10)) +
  labs(x = "K", y = expression("Total W"[k]))

```


#Visualizing clusters

```{r figure19, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(75)
wi_county_km <- kmeans(wi_final_county_factors, centers = 5, nstart = 20)

# add cluster assignments to the data frame
wi_county_clust <- wi_final_county %>%
  mutate(clust3_std = as.character(wi_county_km$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(County=tolower(County))

usa_states <- map_data(map = "county"
                       , region = "wisconsin") %>%
  select(subregion, long, lat, group)

wi_final_cluster <- wi_county_clust %>%
  mutate(County=str_trim(County)) %>%
  left_join(usa_states, by = c("County" = "subregion"))

ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
  geom_polygon(data=wi_final_cluster, aes(x = long, y = lat, group=group, fill=clust3_std), color="black") +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  labs(title = "Clusters in California",
       caption = "*Grey states have no income or unemployment information") +
  theme_void()


```

Cluster characteristics, explain why we chose cluster 5

```{r figure20, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vars_std = c("Avg. Income_std", "Num. Starbucks_std", "Avg. Unemployment Rate_std")

ggpairs(data = wi_county_clust
        , aes(color = clust3_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))

#Cluster 5
wi_poor_county <- wi_county_clust %>%
  filter(clust3_std==5) %>%
  mutate(County=str_to_title(County)) %>%
  inner_join(wi_county_factors, by="County") %>%
  select(County, `Avg. Income`, `Num. Starbucks`,`Avg. Unemployment Rate`)

```


##California

Cluster Analysis
```{r figure21, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

ca_income_city <- city_income %>%
  filter(State_Name=="California") %>%
  group_by(City) %>%
  na.omit() %>%
  summarize("Avg. Income" = mean(Mean, na.rm=TRUE)) 

ca_starbucks_city <- starbucks %>%
  filter(`State/Province`=="CA") %>%
  group_by(`City`) %>%
  summarize("Num. Starbucks" = n())

ca_city_info <- ca_income_city %>%
  left_join(ca_starbucks_city, by = c("City")) %>%
  mutate_all(~replace(., is.na(.), 0))

ca_county_info <- city_income %>%
  filter(State_ab=="CA") %>%
  select(City, County) %>%
  separate(County, into = c("County", "Blank")
           , sep = "County"
           , remove = FALSE) %>% 
  distinct() %>%
  left_join(ca_city_info, by = c("City")) %>%
  group_by(County) %>%
  summarize(
    "Avg. Income" = mean(`Avg. Income`), 
    "Num. Starbucks" = sum(`Num. Starbucks`)
  )

ca_county_factors <- ca_county_info %>%
  left_join(unemployment_county, by = c("County")) %>%
  na.omit()

ca_final_county <- ca_county_factors %>%
  mutate_if(is.numeric, funs(`std`=scale(.) %>% as.vector())) %>%
  select(County, `Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)
  
ca_final_county_factors <- ca_final_county %>%
  select(`Avg. Income_std`, `Num. Starbucks_std`, `Avg. Unemployment Rate_std`)

set.seed(75)
#Create a folder with 10 0s.
fig <- matrix(NA, nrow=10, ncol=2)

#Collecting information for each cluster 
for (i in 1:10){
  fig[i,1] <- i
  fig[i,2] <- kmeans(ca_final_county_factors, centers = i, nstart = 20)$tot.withinss
}

#Visualizing "Elbow Plot"
ggplot(data = as.data.frame(fig), aes(x = V1, y = V2)) +
  geom_point() + 
  geom_line() +
  scale_x_continuous(breaks=c(1:10)) +
  labs(x = "K", y = expression("Total W"[k]))

```


#Visualizing clusters

```{r figure22, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

# apply the k-means algorithm (set the seed to make the results reproducible)
set.seed(75)
ca_county_km <- kmeans(ca_final_county_factors, centers = 4, nstart = 20)

# add cluster assignments to the data frame
ca_county_clust <- ca_final_county %>%
  mutate(clust3_std = as.character(ca_county_km$cluster)) %>%
  arrange(clust3_std) %>%
  mutate(County=tolower(County))

usa_states <- map_data(map = "county"
                       , region = "california") %>%
  select(subregion, long, lat, group)

ca_final_cluster <- ca_county_clust %>%
  mutate(County=str_trim(County)) %>%
  left_join(usa_states, by = c("County" = "subregion"))

ggplot() +
  geom_polygon(data = usa_states, aes(x=long, y = lat, group = group), color="black", fill="black", alpha=0.3) +
  geom_polygon(data=ca_final_cluster, aes(x = long, y = lat, group=group, fill=clust3_std), color="black") +
  scale_size_continuous(range=c(1,12)) +
  scale_color_viridis(trans="log") +
  labs(title = "Clusters in California",
       caption = "*Grey states have no income or unemployment information") +
  theme_void()



```

Cluster characteristics, explain why we chose cluster 1

```{r figure23, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

vars_std = c("Avg. Income_std", "Num. Starbucks_std", "Avg. Unemployment Rate_std")

ggpairs(data = ca_county_clust
        , aes(color = clust3_std) 
        , columns = vars_std
        , upper = list(continuous = "blank"))

ca_poor_county <- ca_county_clust %>%
  filter(clust3_std==1) %>%
  mutate(County=str_to_title(County)) %>%
  inner_join(ca_county_factors, by="County") %>%
  select(County, `Avg. Income`, `Num. Starbucks`,`Avg. Unemployment Rate`)
```


#Final Recommendations

Summarize overall process and thinking and impact of starbucks. Show the data table of all important counties from clusters we chose:


```{r figure24, echo = FALSE, error=FALSE, warning=FALSE, message=FALSE}

final_counties <- ca_poor_county %>%
  bind_rows(nv_poor_county) %>%
  bind_rows(vt_poor_county) %>%
  bind_rows(va_poor_county) %>%
  bind_rows(wi_poor_county) %>%
  filter(`Num. Starbucks`==0)

final_counties

```



